RDF Validation of Metadata \\ on Person-Level and Aggregated Data

### 2015-03-01 paper/paper7/paper7.tex
Searching for metadata and data

\begin{abstract}
For research institutes, data libraries, and data archives,
RDF data validation according to predefined constraints is a much sought-after feature, 
particularly as this is taken for granted in the XML world.
The data most often used in research within the community around research data for the social, behavioural, and economic sciences is person-level data, i.e. data collected about individuals (households, businesses). 
%The DDI-RDF Discovery Vocabulary is used to describe and to discover this kind of research data.
While performing research, the detailed, often access restricted person-level
data is aggregated into less confidential publicly available multi-dimensional tables which answer particular research questions and whose purpose is to gain an interest in further more detailed analyses on the underlying person-level data.
To ensure high quality and trust, metadata and data must satisfy certain criteria - specified in terms of RDF constraints.

@@

\textbf{Metadata on Person-Level Data in RDF.}
For a broader view of the data framework and more detailed analyses we refer to the metadata on person-level data collected for the series \emph{EU-SILC (European Union Statistics on Income and Living Conditions)}\footnote{\url{http://www.gesis.org/missy/eu/metadata/EU-SILC}} 
and publicly provided by the \emph{Microdata Information System (MISSY)}\footnote{\url{http://www.gesis.org/missy/eu/missy-home}}.
\emph{Missy} is an online service platform that provides systematically structured metadata for official statistics on European person-level data sets. This includes data documentation at the study and variable level as well as documentation materials, tools and further information. 
The aggregated variable \emph{formal childcare} is calculated on the basis of six person-level variables like 
\emph{Education at pre-school}\footnote{\url{http://www.gesis.org/missy/eu/metadata/EU-SILC/2011/Cross-sectional/original#2011-Cross-sectional-RL010}}.
%(2) Education at compulsory school,
%(3) Child care at centre-based services,
%(4) Child care at day-care centre,
%(5) Child care by a professional child-minder, and
%(6) Child care by grand-parents, household members, relatives, friends, neighbours.
For each person-level variable detailed metadata is given (definitions and descriptions, theoretical concepts, underlying questions, code lists, frequencies and descriptive statistics, countries, year of data collection, classifications) enabling researchers to replicate the results shown in the aggregated data tables from Eurostat.
Metadata on person-level data is represented in RDF using the \emph{Disco} vocabulary.
The series (\emph{disco:StudyGroup}) \emph{EU-SILC} contains (\emph{disco:inGroup}) one study (\emph{disco:Study}) for each year (\emph{dcterms:temporal}) of data collection, e.g. \emph{EU-SILC 2011}.   
The property \emph{dcterms:spatial} points to the countries (\emph{dcterms:Location} resources which are the same as \emph{GeoNames} resources representing these countries) for which the data has been collected.
The study \emph{EU-SILC 2011} contains (\emph{disco:product}) eight person-level data sets (\emph{disco:LogicalDataSet}).
Data sets include (\emph{disco:variable}) person-level variables (\emph{disco:Variable}) like the six ones needed to calculate the aggregated variable \emph{formal childcare}.
\emph{DCAT} enables to represent aggregated and person-level data inside of data collections like repositories, catalogs, or archives. 
\tb{ToDo: DCAT example}

Metadata on person-level data enables researchers to investigate further research questions based on promising findings of other researchers in form of aggregated data.
One common research question is ,e.g., the comparison of variables like 
\emph{formal childcare} between countries, for which the variable is collected within the context of an individual survey, and other European or non European countries (e.g. OSCE).

The \emph{Simple Knowledge Organization System (SKOS)} is used multiple times within the context of aggregated and person-level (meta)data.
Variables, e.g., are constructed (\emph{disco:representation}) out of values (of one or multiple datatypes) and/or code lists.
The values of the variable \emph{Education at pre-school}, representing the number of education hours during a usual week, are expressed as \emph{skos:Concepts}. 
\emph{Disco} uses \emph{skos:OrderedCollection} to organize them in a particular order in a \emph{skos:memberList}. 
SKOS is also used to form hierarchies of \emph{SBE theoretical concepts} (e.g. Education) with which variables may be associated.
The compete hierarchy (\emph{skos:ConceptScheme}) of theoretical concepts (\emph{skos:Concepts}) of series is built using \emph{skos:narrower}.
The variable \emph{Education at pre-school} is assigned to the theoretical concept \emph{Child Care} which is a narrower concept of \emph{Education} - one of the top concept of the series \emph{EU-SILC}.
Controlled vocabularies serve as extension and reuse mechanism.
For \emph{Disco}, concepts (\emph{skos:Concepts}), organized within controlled vocabularies (\emph{skos:ConceptSchemes}), indicate types of descriptive statistics (\emph{\emph{disco:SummaryStatistics}}) like minimum, maximum, mean, and standard deviation.
From 2012 to 2015, SBE and Linked Data community members developed \emph{XKOS}\footnote{\url{https://github.com/linked-statistics/xkos}} - a SKOS extension to describe formal statistical classifications like the International Standard Classification of Occupations (\emph{ISCO}) and the Statistical Classification of Economic Activities in the European Community \emph{NACE}.

>>>>

\begin{abstract}
For research institutes, data libraries, and data archives,
RDF data validation according to predefined constraints is a much sought-after feature, 
particularly as this is taken for granted in the XML world.
The data most often used in research within the community around research data for the social, behavioural, and economic sciences (\emph{SBE}) is person-level data, i.e. data collected about individuals (households, businesses). 
%The DDI-RDF Discovery Vocabulary is used to describe and to discover this kind of research data.
While performing research, the detailed, often access restricted person-level
data is aggregated into less confidential publicly available multi-dimensional tables which answer particular research questions and whose purpose is to gain an interest in further more detailed analyses on the underlying person-level data.
To ensure high quality and trust, metadata and data must satisfy certain criteria - specified in terms of RDF constraints. 
From 2012 to 2015 together with other Linked Data community members and \emph{SBE} experts, we developed diverse vocabularies to represent \emph{SBE} metadata and data in RDF.

@@

\textbf{Metadata on Person-Level Data in RDF.}
For a broader view of the data framework and more detailed analyses we refer to the metadata on person-level data collected for the series \emph{EU-SILC (European Union Statistics on Income and Living Conditions)}\footnote{\url{http://www.gesis.org/missy/eu/metadata/EU-SILC}} 
and publicly provided by the \emph{Microdata Information System (MISSY)}\footnote{\url{http://www.gesis.org/missy/eu/missy-home}}.
\emph{Missy} is an online service platform that provides systematically structured metadata for official statistics on European person-level data sets. This includes data documentation at the study and variable level as well as documentation materials, tools and further information. 
Aggregated (qb:DataSet) and underlying person-level data sets (\emph{disco:LogicalDataSet}) are connected by \emph{prov:wasDerivedFrom}. 
The aggregated variable \emph{formal childcare} is calculated on the basis of six person-level variables like 
\emph{Education at pre-school}\footnote{\url{http://www.gesis.org/missy/eu/metadata/EU-SILC/2011/Cross-sectional/original#2011-Cross-sectional-RL010}}.
%(2) Education at compulsory school,
%(3) Child care at centre-based services,
%(4) Child care at day-care centre,
%(5) Child care by a professional child-minder, and
%(6) Child care by grand-parents, household members, relatives, friends, neighbours.
For each person-level variable detailed metadata is given (definitions and descriptions, theoretical concepts, underlying questions, code lists, frequencies and descriptive statistics, countries, year of data collection, classifications) enabling researchers to replicate the results shown in the aggregated data tables from Eurostat.
Metadata on person-level data is represented in RDF using the \emph{Disco} vocabulary.
The series (\emph{disco:StudyGroup}) \emph{EU-SILC} contains (\emph{disco:inGroup}) one study (\emph{disco:Study}) for each year (\emph{dcterms:temporal}) of data collection, e.g. \emph{EU-SILC 2011}.   
The property \emph{dcterms:spatial} points to the countries (\emph{dcterms:Location} resources which are the same as \emph{GeoNames} resources representing these countries) for which the data has been collected.
The study \emph{EU-SILC 2011} contains (\emph{disco:product}) eight person-level data sets (\emph{disco:LogicalDataSet}).
Data sets include (\emph{disco:variable}) person-level variables (\emph{disco:Variable}) like the six ones needed to calculate the aggregated variable \emph{formal childcare}.
Metadata on person-level data enables researchers to investigate further research questions based on promising findings of other researchers in form of aggregated data.
One common research question is ,e.g., the comparison of variables like 
\emph{formal childcare} between countries, for which the variable is collected within the context of an individual survey, and other European or non European countries (e.g. OSCE).

\textbf{Organizations, Hierarchies, and Classifications.}
%\textbf{Reusing SKOS.}
The \emph{Simple Knowledge Organization System (SKOS)} is used multiple times within the context of aggregated and person-level (meta)data.
Variables, e.g., are constructed (\emph{disco:representation}) out of values (of one or multiple datatypes) and/or code lists.
The values of the variable \emph{Education at pre-school}, representing the number of education hours during a usual week, are expressed as \emph{skos:Concepts}. 
\emph{Disco} uses \emph{skos:OrderedCollection} to organize them in a particular order in a \emph{skos:memberList}. 
SKOS is also used to form hierarchies of \emph{SBE theoretical concepts} (e.g. Education) with which variables may be associated.
The compete hierarchy (\emph{skos:ConceptScheme}) of theoretical concepts (\emph{skos:Concepts}) of series is built using \emph{skos:narrower}.
The variable \emph{Education at pre-school} is assigned to the theoretical concept \emph{Child Care} which is a narrower concept of \emph{Education} - one of the top concept of the series \emph{EU-SILC}.
Controlled vocabularies serve as extension and reuse mechanism.
For \emph{Disco}, concepts (\emph{skos:Concepts}), organized within controlled vocabularies (\emph{skos:ConceptSchemes}), indicate types of descriptive statistics (\emph{\emph{disco:SummaryStatistics}}) like minimum, maximum, mean, and standard deviation.
From 2012 to 2015, SBE and Linked Data community members developed \emph{XKOS}\footnote{\url{https://github.com/linked-statistics/xkos}} - a SKOS extension to describe formal statistical classifications like the International Standard Classification of Occupations (\emph{ISCO}). 
%and the Statistical Classification of Economic Activities in the European Community \emph{NACE}.

\textbf{Searching for (Meta)data.}
\emph{DCAT} enables to represent aggregated and person-level data inside of data collections like portals, repositories, catalogs, or archives
serving as typical entry points when searching for data.
Users search for aggregated and person-level data records (\emph{dcat:CatalogRecord}) inside data catalogs (\emph{dcat:Catalog}). 
This search differs depending on the users’ information need. 
While it is possible to search for metadata provided inside such a record (e.g. \emph{dcterms:title}, \emph{dcterms:description}), 
users can also formulate more sophisticated queries on aggregated and person-level data sets (\emph{dcat:Dataset}) or their
distributions (\emph{dcat:Distribution}), which are part of the records. 
Users may want to search for data sets covering particular topical (\emph{dcat:keyword}, \emph{dcat:theme}), temporal (\emph{dcterms:temporal}),  or  spatial  coverages (\emph{dcterms:spatial}), 
or certain formats in which the data distribution is available (\emph{dcterms:format}).
